---
title: "Master_thesis"
author: "Visweshwar"
date: "05/08/2023"
output: html_document
---
```{r, warning=FALSE}
## Setting up all the libraries that were used 
library(lidR)
library(raster)
library(dplyr)
library(sf)
library(caret)
library(ggplot2)
library(fields) # image.plot
library(sp)
library(data.table)
library("openair")
library(scales)
```

LIDAR Data Processing

```{r, warning=FALSE}
## Reading in the Lidar las file
amtsvenn_catalog <- readLAS("D:/Vishwa/lidars/terra_las/cloude0cf101272f2bc76.las")
amtsvenn_catalog <- filter_poi(amtsvenn_catalog, Z < 115)

## Reading the Ec tower location point and setting the CRS
ec_tower <- st_read("D:/Vishwa/r works/ec_tower.shp")
ec_tower <- st_transform(ec_tower, "+proj=utm +zone=32 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs +type=crs")

## Generating 100m buffer around the EC tower to be used as the extent
buffer <- st_buffer(ec_tower, 100)

## Performing coordinate transformation of the lidar data in order to match the buffer and clip the lidar data within the buffer extent
transformed_pc <- st_transform(amtsvenn_catalog, st_crs(buffer))
clipped_amtsvenn <- clip_roi(transformed_pc, st_bbox(buffer))

##Visualizing the mean values within 1m resolution
amts_elevation <- pixel_metrics(clipped_amtsvenn, ~mean(Z), 1)
amts_elevation <- raster(amts_elevation)
plot(amts_elevation)

## Ground classification of the lidar data using the Cloth simulation filter algorithm
classified_peat <- classify_ground(transformed_pc, algorithm = csf(class_threshold = 0.4))

```

Processing of DSM and DTM in order to obtain the roughness length values

```{r, warning=FALSE}
##Generating DSM and DTM from the LIDAR data and clipping it to the buffer extent
dtm_tin <- rasterize_terrain(classified_peat, res = 1, algorithm = tin())
dtm_tin<- raster(dtm_tin)
dtm_tin <- crop(dtm_tin, amts_elevation)
plot(dtm_tin, main= "Digital Terrain Model")

dsm_tin <- rasterize_canopy(classified_peat, res = 1, algorithm =  pitfree())
dsm_tin <- raster(dsm_tin)
dsm_tin <- crop(dsm_tin, amts_elevation)
plot(dsm_tin, main= "Digital Surface Model")

## Normalizing DSM and DTM values based on the minimum value from DTM
norm_dsm <- dsm_tin-minValue(dtm_tin)
norm_dtm <- dtm_tin-minValue(dtm_tin)

## Calculating roughness length from DSM and DTM
z0_raster <- norm_dsm*0.1
z0_matrix <- as.matrix(z0_raster)

plot(z0_raster, main="Roughness length values from DSM")

## Since the FFP model doesn't accept 0 value, 0.01 was replaced at the location of the minimum value
norm_dtm[norm_dtm==0] <-0.01
dtm_z0_raster <- norm_dtm*0.1
dtm_z0 <- as.matrix(dtm_z0_raster)
plot(dtm_z0_raster, main="Roughness length values from DTM")

```
Cleaning Eddy Covariance data

```{r, warning=FALSE}
## Reading in EC Tower flux data
ffp.data <- read.csv("D:/Vishwa/r works/flux_data.csv")
ffp.data <- ffp.data[,-1]


## omit NAs in the data, so that the calculation always works
final_dat <- na.omit(ffp.data)

##Calculating planetary boundary layer height based on appendix B of Kljun et al. (2015)
lat <- 52.1757  # latitude of the station
rot <- 7.2921*10^(-5) # angular velocity of the earth
omega <- 2*pi/(23*3600 + 56*60 + 4.1) # Rotationsgeschwindigkeit, same as above
f <- 2*rot*sin((lat*pi/180))# Coriolis parameter

## Adding PBL height to the EC tower data
final_dat$h <- final_dat$L/3.8*(-1+(1+2.28*final_dat$u_star/(f*final_dat$L))^0.5)

## Retrieving good quality data and removing any na values
final_dat <- final_dat[final_dat$qc_co2_flux==0,]
final_dat <- na.omit(final_dat)
head(final_dat)

```
Calculating Footprint Climatology
```{r, warning=FALSE}
## Modifiying the Matrix roughness lengths to fit the model grid
z0_matrix  <- z0_matrix[nrow(z0_matrix):1, ]
dtm_z0  <- dtm_z0[nrow(dtm_z0):1, ]

## Assigning model parameters and data
dat <- final_dat
day_dat <- dat[dat$HH_CET > 9 & dat$HH_CET < 17, ] # Flux data measured during the day between 10 am to 4pm 
nig_dat <- dat[dat$HH_CET < 7, ]  # Flux data measured during the night between 12 am to 6am

d.xy <- 1 # pixel resolution in m
dd <-c(-100, 100, -100, 100) # footprint grid

## Wind rose plot of the EC tower data during November to Feb
windRose(dat, ws="u_mean", wd="wind_dir", paddle=F, cols=c("gray", "orange", "red"), angle=15, ws.int=2, par.settings=list(fontsize=list(text=24)), key=list(labels=c("<2", "2-4", "4-6", ">6")))

```
Footprint climatology based on the original FFP (scalar roughness length value) is calculated first. Source area footprints up to 70% are computed.

```{r, warning=FALSE}

setwd("D:/Vishwa/r works/")
source('user_defined_function.R') ## User defined functions for the purpose of analysis
source('original_ffp.R') ##Original FFP model

## Footprint climatology for the entire duration 
(nvals <- nrow(dat))
output_ori <- calc_footprint_FFP_climatology(dat$zm, dat$z0, dat$u_mean, dat$h, dat$L,
                                             dat$sigma_v, dat$u_star, dat$wind_dir,
                                             domain = dd, dx = 1, dy = 1,
                                             nx = NULL, ny = NULL, r= seq(10,70,10), rslayer = NULL, smooth_data = 0,
                                             crop = NULL, pulse = 1, fig = 1)


## Footprint climatology of EC data during daytime
(nvals <- nrow(day_dat))

day_output_ori <- calc_footprint_FFP_climatology(day_dat$zm, day_dat$z0, day_dat$u_mean, day_dat$h, day_dat$L,
                                                 day_dat$sigma_v, day_dat$u_star, day_dat$wind_dir,
                                                 domain = dd, dx = 1, dy = 1,
                                                 nx = NULL, ny = NULL, r= seq(10,70,10), rslayer = NULL, smooth_data = 0,
                                                 crop = NULL, pulse = 1, fig = 1)


## Footprint climatology of EC data during nighttime
(nvals <- nrow(nig_dat))

nig_output_ori <- calc_footprint_FFP_climatology(nig_dat$zm, nig_dat$z0, nig_dat$u_mean, nig_dat$h, nig_dat$L,
                                                 nig_dat$sigma_v, nig_dat$u_star, nig_dat$wind_dir,
                                                 domain = dd, dx = 1, dy = 1,
                                                 nx = NULL, ny = NULL, r= seq(10,70,10), rslayer = NULL, smooth_data = 0,
                                                 crop = NULL, pulse = 1, fig = 1)

## User defined function (total_day_night) which converts the coordinate system (CRS) of the obtained footprint climatology from a local CRS to a global CRS and calculates the area for each source percent contours. 
original <- total_day_night(output_ori, day_output_ori, nig_output_ori)
original_poly <- original[[1]]
original_poly$model <- "original"
original_raster <- original[[2]]

plot(st_geometry(original_poly[1:7,]), main="Original FFP model source area footprints upto 70%")
```
Once original FFP footprints are calculated, DSM based footprint is computed next.

```{r, warning=FALSE}
source('spatial_FFP.R')## Modified FFP model

## DSM based footprint climatology
(nvals <- nrow(dat))
output <- calc_footprint_FFP_climatology(dat$zm, z0_matrix, dat$u_mean, dat$h, dat$L,
                                         dat$sigma_v, dat$u_star, dat$wind_dir,
                                         domain = dd, dx = 1, dy = 1,
                                         nx = NULL, ny = NULL, r= seq(10,70,10), rslayer = NULL,  pulse = 1, fig = 1)

## Footprint climatology of EC data during daytime
(nvals <- nrow(day_dat))

day_output <- calc_footprint_FFP_climatology(day_dat$zm, z0_matrix, day_dat$u_mean, day_dat$h, day_dat$L,
                                             day_dat$sigma_v, day_dat$u_star, day_dat$wind_dir,
                                             domain = dd, dx = 1, dy = 1,
                                             nx = NULL, ny = NULL, r= seq(10,70,10), rslayer = NULL, pulse = 1, fig = 1)

## Footprint climatology of EC data during nighttime
(nvals <- nrow(nig_dat))

nig_output <- calc_footprint_FFP_climatology(nig_dat$zm, z0_matrix, nig_dat$u_mean, nig_dat$h, nig_dat$L,
                                             nig_dat$sigma_v, nig_dat$u_star, nig_dat$wind_dir,
                                             domain = dd, dx = 1, dy = 1,
                                             nx = NULL, ny = NULL, r= seq(10,70,10), rslayer = NULL, pulse = 1, fig = 1)


## User defined function (total_day_night) which converts the coordinate system (CRS) of the obtained footprint climatology from a local CRS to a global CRS and calculates the area for each source percent contours. 
modified <- total_day_night(output, day_output, nig_output)
modified_poly <- modified[[1]]
modified_poly$model <- "dsm"
modified_poly <- na.omit(modified_poly)
modified_raster <- modified[[2]]
plot(modified_poly)

```
DTM based footprint is computed in this section.
```{r setup, warning=FALSE}

## DTM based footprint climatology
(nvals <- nrow(dat))
output_dtm <- calc_footprint_FFP_climatology(dat$zm, dtm_z0, dat$u_mean, dat$h, dat$L,
                                         dat$sigma_v, dat$u_star, dat$wind_dir,
                                         domain = dd, dx = 1, dy = 1,
                                         nx = NULL, ny = NULL, r= seq(10,70,10), rslayer = NULL, smooth_data = 0,
                                         crop = NULL, pulse = 1, fig = 1)

## Footprint climatology of EC data during daytime
(nvals <- nrow(day_dat))

day_output_dtm <- calc_footprint_FFP_climatology(day_dat$zm, dtm_z0, day_dat$u_mean, day_dat$h, day_dat$L,
                                             day_dat$sigma_v, day_dat$u_star, day_dat$wind_dir,
                                             domain = dd, dx = 1, dy = 1,
                                             nx = NULL, ny = NULL, r= seq(10,70,10), rslayer = NULL, smooth_data = 0,
                                             crop = NULL, pulse = 1, fig = 1)

## Footprint climatology of EC data during nighttime
(nvals <- nrow(nig_dat))

nig_output_dtm <- calc_footprint_FFP_climatology(nig_dat$zm, dtm_z0, nig_dat$u_mean, nig_dat$h, nig_dat$L,
                                             nig_dat$sigma_v, nig_dat$u_star, nig_dat$wind_dir,
                                             domain = dd, dx = 1, dy = 1,
                                             nx = NULL, ny = NULL, r= seq(10,70,10), rslayer = NULL, smooth_data = 0,
                                             crop = NULL, pulse = 1, fig = 1)


## User defined function (total_day_night) which converts the coordinate system (CRS) of the obtained footprint climatology from a local CRS to a global CRS and calculates the area for each source percent contours. 
dtm_modified <- total_day_night(output_dtm, day_output_dtm, nig_output_dtm)
dtm_modified_poly <- dtm_modified[[1]]
dtm_modified_poly$model <- "dtm"
dtm_modified_poly <- na.omit(dtm_modified_poly)

dtm_modified_raster <- dtm_modified[[2]]
plot(st_geometry(dtm_modified_poly))
```
```{r, warning=FALSE}
## Plotting the cumulative source area percent of the different models
polygon_area <- rbind(original_poly, modified_poly, dtm_modified_poly)
polygon_area$sa_perc <- as.character(polygon_area$sa_perc)
my_sum1 <- polygon_area %>%
  group_by(sa_perc, model) %>%
  summarise( 
    n=n(),
    mean=as.integer(mean(area)),
    sd=sd(area)
  ) 


ggplot(my_sum1, aes(x=sa_perc, y=mean, fill=model)) + 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                position=position_dodge(.9)) + theme_bw()+
  theme(axis.text = element_text(size = 12),axis.title = element_text(size = 18),
        legend.text = element_text(size = 13),legend.title = element_text(size = 18),plot.title = element_text(size=20) ) + 
  labs(title = "Cumulative area of different source percent contour using different roughness based models", x = "percentage source", y = "Cumulative Area (m2)")

```
```{r, warning=FALSE}

average_size_decrease <- data.frame(sa_perc= original_poly$sa_perc, original = original_poly$area,
                                    dsm = modified_poly$area, dtm= dtm_modified_poly$area)

average_size_decrease$dsm_ori <- ((average_size_decrease$dsm/ average_size_decrease$original)*100)
average_size_decrease$dtm_ori <- ((average_size_decrease$dtm/ average_size_decrease$original)*100)

#area agreement between dsm and original model (subtracting this value from 100 gives the decrease in area)
mean(average_size_decrease$dsm_ori)
#area agreement between dtm and original model(subtracting this value from 100 gives the decrease in area)
mean(average_size_decrease$dtm_ori)

```


Preparing predictors for the Random forest algorithm (Multi spectral and lidar DSM)

```{r, warning=FALSE}
### Reading in Multi spectral data
multi_spec_amtsvenn <- stack("D:/Vishwa/Amtsvenn_Micasense_25102022.tif")

### Creating high resolution DSM from Lidar data
class_dsm_tin <- rasterize_canopy(classified_peat, res = 0.0739835, algorithm =  pitfree())
class_dsm_tin <- raster(class_dsm_tin)

### Filling in gaps using a moving window of 3x3 and matching multispectral data with lidar data CRS
fill.na <- function(x, i=5) { if (is.na(x)[i]) { return(mean(x, na.rm = TRUE)) } else { return(x[i]) }}
w <- matrix(1, 3, 3)
class_dsm_tin <- terra::focal(class_dsm_tin, w, fun = fill.na)
crs(multi_spec_amtsvenn) <- crs(class_dsm_tin) <- CRS('+init=EPSG:25832')

### Field polygon to be used for cropping and masking 
field_poly <-read_sf("D:/Vishwa/r works/field_poly.shp")

### Calculating vegetation indices
ndvi_am <- (multi_spec_amtsvenn[[5]] - multi_spec_amtsvenn[[3]]) / (multi_spec_amtsvenn[[5]] + multi_spec_amtsvenn[[3]]) 
chloro_am <- (multi_spec_amtsvenn[[5]] / multi_spec_amtsvenn[[1]]) - 1
ngrdi_am <-  (multi_spec_amtsvenn[[1]] - multi_spec_amtsvenn[[3]]) / (multi_spec_amtsvenn[[1]] + multi_spec_amtsvenn[[3]])

### Stacking all the predictors variables
total_rasters <- stack(multi_spec_amtsvenn, ndvi_am, chloro_am, ngrdi_am)
total_rasters <- mask(total_rasters, field_poly)
total_rasters <- crop(total_rasters, field_poly)

####### Resampling lidar DSM to match the multi spectral data
class_dsm_tin <- mask(class_dsm_tin, field_poly)
class_dsm_tin <- crop(class_dsm_tin, field_poly)
resample_dsm <- resample(class_dsm_tin, total_rasters)

## total stack of the rasters used for predictions
sta <- stack(total_rasters, resample_dsm)

names_multi <- c("green", "blue", "red", "red_edge", "NIR", "ndvi", "CI", "ngrdi", "dsm")
names(sta) <- names_multi
plot(sta)

```

```{r, warning=FALSE}
## reading in training and test data for the RF classification
trainDat <-as.data.frame(read_sf("D:/Vishwa/r works/rf_training.shp"))
testDat <- as.data.frame(read_sf("D:/Vishwa/r works/rf_test.shp"))

nrow(trainDat)
response <- "MC_name"   #prediction variable
predictors <- names_multi # predictors
set.seed(100)
print("Creating the model")

model <- train(trainDat[,predictors],trainDat[,response],method="rf",trControl=trainControl(method="cv"),importance=TRUE) #model training

prediction <- predict(sta, model) #predicted LULC classification
plot(prediction, col.regions=c("red","darkgreen","yellow","green","blue","brown"))


pred_valid <- predict(model,testDat) #accuracy assessments
confusion_mat<- confusionMatrix(pred_valid, as.factor(testDat$MC_name)) # confusion matrix

ggplotConfusionMatrix(confusion_mat) #confusion matrix plot
```

Classification based plots

```{r, warning=FALSE}
## Converting the classification image to a dataframe for plotting purposes
pred_df <- as.data.frame((prediction), xy=TRUE)%>%
  na.omit() #

##ggplot of the Microform classification image 
ggplot(data = pred_df) +
  geom_raster(aes(x = x, 
                  y = y, 
                  fill = (layer_value))) + 
  scale_fill_manual(name = "Microform",
                    values = c("yellow","red","green","pink","brown", "darkgreen"),
                    na.translate = FALSE) + theme_bw()+
  coord_sf(expand = FALSE) + ggtitle("Microform cover at Amtsvenn")+theme(plot.title = element_text(size=20) ) 

##Field level microform coverage plot
field_lvl <- freq(prediction, useNA='no') #counting the frequency of the different microforms in the prediction image
apc <- prod(res(prediction)) # area of per pixel
field_lvl <- as.data.frame(cbind(field_lvl, area=field_lvl[,2] * apc))# Area coverage of each microform
colnames(field_lvl) <- c("Microform", "count", "area")
field_lvl$Microform <- prediction@data@attributes[[1]][["value"]]

## Barplot of Microform cover area at the field level
ggplot(data = field_lvl, aes(x = Microform, 
                             y = area,  fill=Microform))+ geom_bar(stat="identity")+ theme_bw()+ 
    geom_text(aes(label= round (area)), position=position_dodge(width=0.9), vjust=-0.25)+
    theme(axis.text = element_text(size = 16),axis.title = element_text(size = 18),
          legend.text = element_text(size = 18),legend.title = element_text(size = 18),plot.title = element_text(size=20) ) + labs(title = "Microform cover by area (Field level)", x = "Microforms", y = bquote("Area" ~(m^2)))

```

Wind direction Analysis
```{r, warning=FALSE}
## reading in the wind direction buffer created from QGIS. For each wind sector, the shapefile has its representative direction
wind_direction_buffer <- st_read("D:/Vishwa/r works/wind_direction.shp")

buffer_55 <- st_buffer(ec_tower, 55)#creating a 55m (based on distance from tower to the boundary of the contour) buffer from the EC tower 
buffer_55 <- st_transform(buffer_55, st_crs(wind_direction_buffer))# Matching the CRS both polygons
wind_direction_buffer <- (st_intersection(wind_direction_buffer, buffer_55)) # Cropping the wind direction polygon to the 55m buffer distance as the original polygon had a distance of 120m

## Calculating the within polygon cover of each microform along each direction
dom_lulc_wd <-extract(prediction, wind_direction_buffer, na.rm= T) #retrieves pixel value from within each wind direction

#cleaning and preparing the data to obtain a dataframe with desired format
names(dom_lulc_wd) <- wind_direction_buffer$direction
dom_lulc_wd <-lapply(dom_lulc_wd, function(x)sort(table(x),decreasing=TRUE)[1:6])
dom_lulc_wd<- lapply(dom_lulc_wd, as.data.frame)
dom_lulc_wd <- dplyr::bind_rows(dom_lulc_wd, .id = "variable")
dom_lulc_wd$x <- as.character(dom_lulc_wd$x) #to change the variable from a factor to integer it needs to be changed to character first
dom_lulc_wd$x <- as.integer(dom_lulc_wd$x)#change it to integer
dom_lulc_wd$area <- dom_lulc_wd$Freq*apc 
cover <- prediction@data@attributes[[1]][["value"]]

## Assigning the cover name to their code name. (Raster uses factors to assign name of the cover to the codes. When changing the raster to a dataframe, only the codes are retrieved.)
for(i in 1:6){
  dom_lulc_wd$x[dom_lulc_wd$x==i] <- cover[i]
}

##Barplot comparison of microform cover along different wind direction
ggplot(dom_lulc_wd, aes(x=variable, y=area, fill=x)) + 
  geom_bar(stat="identity", color="black") + 
  theme(axis.text = element_text(size = 12),axis.title = element_text(size = 18),
        legend.text = element_text(size = 13),legend.title = element_text(size = 18),plot.title = element_text(size=20) ) + 
  labs(title = "Dominant microform cover in different wind directions", x = "Wind directions", y = "Area (sq.m)", fill ="microform")

```

Footprint level Microform coverage. 

```{r, warning=FALSE}
#Data preparation. User defined function footpr_level will extract the pixels from the prediction image within the footprints and calculates the area for each cover that was retrieved
fp_cl_area_day <- footpr_level(list(original_poly[original_poly$tperiod=="day" ,], modified_poly[modified_poly$tperiod=="day",], dtm_modified_poly[dtm_modified_poly$tperiod=="day",]), prediction, apc)# Microform cover within day footprints

fp_cl_area_nigh <- footpr_level(list(original_poly[original_poly$tperiod=="night",], modified_poly[modified_poly$tperiod=="night",], dtm_modified_poly[dtm_modified_poly$tperiod=="night",]), prediction, apc)# Microform cover within night footprints

fp_cl_area_all <- footpr_level(list(original_poly[original_poly$tperiod=="all",], modified_poly[modified_poly$tperiod=="all" ,], dtm_modified_poly[dtm_modified_poly$tperiod=="all",]), prediction, apc)# Microform cover within the total data footprints

fp_cl_area_day$tperiod <- "day"
fp_cl_area_all$tperiod <- "all"
fp_cl_area_nigh$tperiod <- "night"
total_fp_class <- rbind(fp_cl_area_day, fp_cl_area_all, fp_cl_area_nigh)

#Calculating the mean and std dev variation of the area of the different microforms for the different models 
my_sum <- total_fp_class %>%
  group_by(microform, model) %>%
  summarise( 
    n=n(),
    mean=mean(area),
    sd=sd(area)
  ) 

## Barplot of the Microform cover found within the footprints
ggplot(my_sum, aes(x=microform, y=mean, fill=model)) + 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2,
                position=position_dodge(.9))  + 
  labs(title = "Microform cover found within 70% source area", x = "Microforms", y = bquote("Area" ~(m^2))) + theme_bw()+
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 14),
        legend.text = element_text(size = 14),legend.title = element_text(size = 18),plot.title = element_text(size=20) )

```
Calculating the distance of the peak value from the EC tower
```{r, warning=FALSE}
## User defined function peak_location is used to estimate the distance of the peak value from the EC tower
original_peak<- lapply(original_raster, peak_location, ec_tower) %>% unlist()
dsm_peak <- lapply(modified_raster, peak_location, ec_tower) %>% unlist()
dtm_peak <- lapply(dtm_modified_raster, peak_location, ec_tower) %>% unlist()
peak_location <- data.frame(original=original_peak, dsm= dsm_peak, dtm= dtm_peak)
head(peak_location)
```

